Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-02-13 paperIEEE.tex
minor fixes and fixmes.
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@129 db7c848b-7755-4df2-b763-8796797afa52

\textsf{Equalizer} itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}; load-balancing resources for multi-display
installations~\cite{EEP:11}; asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}; as well as work queueing~\cite{SPEP:16}.
Additionally, complex large scale and out-of-core multiresolution
rendering approaches have been parallelized and implemented with
Equalizer~\cite{GMBP:10, GEMPG:13}, demonstrating the feasibility of
the framework to be used with complex rendering algorithms and 3D model
representations.

@@

Sort-first (\fig{fig:loadeq}) and sort-last load balancing are the most obvious
optimizations for these parallel rendering modes. Our load equalizers are fully
transparent for application developers; that is, they use a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherency. Equalizer implements two different algorithms, a
\textsf{load\_equalizer} and a \textsf{tree\_equalizer}. The result section
provides some evidence on the strengths and weaknesses of both algorithms.

@@

The package equalizers implement client-affinity models for tile and chunk
compounds (\sref{sec:tile}). A \textsf{tile\_equalizer} and
\textsf{chunk\_equalizer} create the packages and change the assignment of them
to individual nodes, based on an affinity model specified in the equalizer.
In~\cite{SPEP:16}, we explore this approach in detail.

@@

The DFR equalizer (\fig{fig:dfr}) provides a functionality similar to dynamic
video resizing \cite{MBDM:97}, that is, it maintains a constant framerate by
adapting the rendering resolution of a fill-limited application. In
\textsf{Equalizer}, this works by rendering into a source channel (typically on
a FBO) separate to the destination channel, and then scaling the rendering
during the transfer (typically through an on-GPU texture) to the destination
channel. The DFR equalizer monitors the rendering performance and accordingly
adapts the resolution of the source channel and zoom factor for the source to
destination transfer. If the performance and source channel resolutions allows,
this will not only subsample, but also supersample the destination channel to
reduce aliasing artefacts.

@@

Asynchronous compositing pipelines rendering with compositing operations, by
executing the image readback, network transfer and image assembly from threads
running in parallel to the rendering threads. In \cite{EBAHMP:12}, we provide
the details of the implementation and experimental data showing an improvement
of the rendering performance of over 25\% for large node counts.

@@

Compression for the compositing step is critical for performance. This not only
applies to the well-researched network transfer step, but also for the transfer
between GPU and CPU. \textsf{Equalizer} supports a variety of compression
algorithms, from very fast RLE encoding, YUV subsampling on the GPU to JPEG
compression. These algorithms are implemented as runtime-loaded plugins,
allowing easy extension and customization to application-specific compression.
In \cite{MEP:10}, we show this to be a critical step for interactive performance
at scale.

>>>>

\textsf{Equalizer} itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}; load-balancing resources for multi-display
installations~\cite{EEP:11}; asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}; as well as work queueing~\cite{SPEP:16}. \FIXME{Why ; instead of , ?}
Additionally, complex large scale and out-of-core multiresolution
rendering approaches have been parallelized and implemented with
Equalizer~\cite{GMBP:10, GEMPG:13}, demonstrating the feasibility of
the framework to be used with complex rendering algorithms and 3D model
representations.

@@

Sort-first (\fig{fig:loadeq}) and sort-last load balancing are the most obvious
optimizations for these parallel rendering modes. Our load equalizers are fully
transparent for application developers; that is, they use a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherency \FIXME{Shouldn't that be frame-to-frame {\em coherence}?}. Equalizer implements two different algorithms, a
\textsf{load\_equalizer} and a \textsf{tree\_equalizer}. The result section
provides some evidence on the strengths and weaknesses of both algorithms.

@@

The package equalizers implement client-affinity models for tile and chunk
compounds (\sref{sec:tile}). A \textsf{tile\_equalizer} or
\textsf{chunk\_equalizer} creates the packages and changes the assignment of them
to individual nodes, based on an affinity model specified in the equalizer.
In~\cite{SPEP:16}, we explore this approach in detail.

@@

The DFR equalizer (\fig{fig:dfr}) provides a functionality similar to dynamic
video resizing \cite{MBDM:97}, that is, it maintains a constant framerate by
adapting the rendering resolution of a fill-limited application. In
\textsf{Equalizer}, this works by rendering into a source channel (typically on
a FBO) separate to the destination channel, and then scaling the rendering
during the transfer (typically through an on-GPU texture) to the destination
channel. The DFR equalizer monitors the rendering performance and accordingly
adapts the resolution of the source channel and zoom factor for the source to
destination transfer. If the performance and source channel resolutions allow,
this will not only subsample, but also supersample the destination channel to
reduce aliasing artefacts.

@@

\FIXME{Sentence incomplete.}
Asynchronous compositing pipelines rendering with compositing operations, by
executing the image readback, network transfer and image assembly from threads
running in parallel to the rendering threads. In \cite{EBAHMP:12}, we provide
the details of the implementation and experimental data showing an improvement
of the rendering performance of over 25\% for large node counts.

@@

Compression for the compositing step is critical for performance. This not only
applies to the well-researched network transfer step, but also for the transfer
between GPU and CPU. \textsf{Equalizer} supports a variety of compression
algorithms, from very fast run-length encoding and YUV subsampling on the GPU to JPEG
compression. These algorithms are implemented as runtime-loaded plugins,
allowing easy extension and customization to application-specific compression.
In \cite{MEP:10}, we show this to be a critical step for interactive performance
at scale.
